---
title: "DataMoves"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DataMoves}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(KCundercover)
library(tidyverse)
neardeath <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2026/2026-07-21/nde_experiences.csv")
neardeath1 <- betterthanfactoring(neardeath, c("gender", "classification", "country", "category", "language", "ai_obe", "ai_unity", "ai_hellish", "ai_clinical", "ai_esp", "ai_past_lives", "ai_world_future", "ai_aliens"))
```

# Data Set Used throughout KCundercover Data Moves:

The data set used to make create the example data moves contained with in this vignette is 'neardeath1'. This data set contains a collection of 18 variable that can be used to assess and compare a near death experience, such as gender, country, and greyson_score. Greyson_score is a score under to assess whether or not a near death experience in genuine or not, any greyson_score of 7 or higher it considered a genuine near death experience. This data set contains 589 near death experience entries. All entries come from a different near death experience.

# Data Moves:

## Filter:

The filter data move in this example filters the near death experience data set called "neardeath1" to responses that were submitted in English and by females. The top six rows of this column were then outputted by the head() function. 

```{r}
neardeath1 %>%
  filter(language == "english", gender == "Female") %>% 
  head()
```

## Group By/Summarize:

The group by/summaries data move in this example groups the near death experience data set called "neardeath1" by the two different categories of gender responses. Then the responses are summarized to output the mean greyson score for both genders. 

```{r}
neardeath1 %>%
  group_by(gender) %>%
  summarize(mean_greyson_score = mean(greyson_score, na.rm = TRUE))
```

## Join:

The join data move in this example groups the near death experience data set called "neardeath1" by the two different categories of gender responses. Then the responses are summarized to output the mean narrative length for both genders into a new variable called "average_narrative_length" which is saved into the new data set created called "totals". 

```{r}
totals <- neardeath1 %>%
  group_by(gender) %>%
  summarise(average_narrative_length = mean(narrative_length))

neardeath1 <- neardeath1 %>%
  left_join(totals, by = c("gender"))

neardeath1 <- neardeath1 %>%
  mutate(narrativelength_difference_from_average = narrative_length - average_narrative_length)
```

## Arrange:

The first section of code shows uses the death near data set, called "neardeath1", and the head function to displays the first 6 rows in default order of the data set for reference. Then the arrange data move in the second section of code takes the near death experience data set and sorts the data in ascending order using the arrange function by the variable entry_id. Lastly, the head function returns the first 6 rows of the data set so the arranged order is visible. 

```{r}
neardeath1 %>% 
  head()

neardeath1 %>%
  arrange(entry_id) %>%
  head()
```

## Mutate:

The mutate data move in this section of code takes the near death experience data set, called "neardeath1", and uses the mutation function to create a new column, called "greyson_score_geniune" within the near death experience data set. The new variable categories the variable greyson_score into two categories, "Ungeniune NDE" and "Geniune NDE".

```{r}
neardeath1 %>%
  mutate(greyson_score_geniune = case_when(
    greyson_score > 7 ~ "Ungeniune NDE",
    greyson_score <= 7 ~ "Geniune NDE"
  )) %>%
  head()
```

# Betterthanfactoring Helping Function 

The help function contained within the KCundercover package is called 'betterthanfactoring'. The goal of the betterthanfactoring helper function is to make the data cleaning process easier and less time consuming allowing users to shift their focus to their data analysis. The betterthanfactoring function allows users to tell the function a data set then a list of concatenated variables that need to be factored. The betterthanfactoring function will then factor the list of variables for the user, allowing the new factored variable to saved back into the original data set. The inter workings of the betterthanfactoring function can be seen in the functions help document with a description of the inputs and outputs of the function. An example of using the betterthanfactoring helper function in use can be seen below.

```{r}
class(neardeath$gender)
test <- betterthanfactoring(neardeath, c("gender", "country"))
class(test$gender)
```

